### 2021-04-18 phd/thesis proposal/Thesis_proposal.tex

\documentclass[12pt, twoside]{book} % 12 pt font, two-sided book style
\usepackage[a4paper, includehead, headheight=0.6cm, inner=3cm ,outer=2.5cm, top=2.5 cm, bottom=2.5cm]{geometry}  % Changing size of document
\usepackage[english]{babel} % The document is in English
\usepackage[utf8]{inputenc} % UTF8 encoding
\usepackage[T1]{fontenc} % Font encoding

\usepackage{graphicx} % For including images
\graphicspath{{./Images/}} % Specifies the directory where pictures are stored
\usepackage{longtable} % tables that can span several pages
\usepackage[bf]{caption} % caption: FIG in bold
\usepackage{fancyhdr} % For the headers

\newcommand{\numberedchapter}{ % Preparation for numbered chapters
	\cleardoublepage % To make sure the previous headers are passed
	\fancyhead[RE]{{\bfseries \leftmark}}% Headers for left pages
	\fancyhead[LO]{{\bfseries \rightmark}}}% Headers for right pages
\newcommand{\unnumberedchapter}[1]{ % Preparation for unnumbered chapters
	\cleardoublepage % To make sure the previous headers are passed
	\addcontentsline{toc}{chapter}{#1} % Also adds the chapter name to the Contents
	\fancyhead[RE]{{\bfseries #1}} % Headers for left pages
	\fancyhead[LO]{}}%Headers for right pages

\usepackage{emptypage} % No headers on an empty page

\usepackage{eso-pic} % For the background picture on the title page

\usepackage{hyperref} % Adds clickable links at references
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{array}
\usepackage{colortbl}

%----------------------------------------------------------------------------------------
%	ADD YOUR CUSTOM VALUES, COMMANDS AND PACKAGES
%----------------------------------------------------------------------------------------

% Open Preamble/mydefinitions.tex and enter some values (name, thesis title...) 
% and include your own custom LaTeX functions and packages

\input{Preamble/mydefinitions}

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\pagestyle{empty} % No page numbers
\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the preamble pages

\begin{titlepage}
\begin{center}
\vfill
{\large \scshape Pontifical Catholic University of Peru\\Postgraduate School}\\[1.4cm]
{\Large Doctoral Thesis Proposal}\\[0.5cm]
\rule{\textwidth}{1.5pt}\\[0cm]
{\huge \bfseries \thesistitle \par \ }\\[-0.5cm]
\rule{\textwidth}{1.5pt}\\[2.5cm]
by\\[1cm]
{\large \bfseries\name}\\[2cm]
Supervisor\\[1cm]
{\large \textbf{\supervisor}} \\ 
\vfill
%\ifx\cosupervisor\undefined\else{\hfill \large Co-Supervisor: \textbf{\cosupervisor}} \\ \fi%
\vspace{1cm}
\submissiondate
\end{center}
\end{titlepage}

%----------------------------------------------------------------------------------------
%	PREAMBLE PAGES (comment out unnecessary pages)
%----------------------------------------------------------------------------------------

\pagestyle{fancy} % Changes the headers
\renewcommand{\chaptermark}[1]{ \markboth{#1}{}} % Getting the chapter name right
\renewcommand{\sectionmark}[1]{\markright{\thesection\; #1}} % Getting the section name right
\fancyhf{}% Clears header and footer
\fancyhead[RO,LE]{\thepage} % page number on the outside of headers

\input{Preamble/abstract}

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES
%----------------------------------------------------------------------------------------

\unnumberedchapter{Contents}
\tableofcontents % Write out the Table of Contents

%----------------------------------------------------------------------------------------
%	THESIS MAIN TEXT - CHAPTERS
%----------------------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\mainmatter % Begin numeric (1,2,3...) page numbering

\numberedchapter
\input{MainText/chapter1} % Import your chapters here
\input{MainText/chapter2}
\pagestyle{plain}
\input{MainText/chapter3} 
%\input{MainText/chapter4} 
%\input{MainText/chapter5}

%----------------------------------------------------------------------------------------
%	APPENDICES
%----------------------------------------------------------------------------------------

%\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
%\appendix % Starts of appendices
%
%\numberedchapter
%\input{MainText/appendixA}
%\input{MainText/appendixB}
%\input{MainText/appendixC}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\unnumberedchapter{Bibliography} % Title of the unnumbered chapter
\bibliography{Preamble/Thesis_bibliography} % The references information are stored in the file named "Thesis_bibliography.bib"

\end{document}