### 2021-05-09 phd/thesis plan/MainText/chapter3.tex

\chapter{Figures, tables and images} \label{chap-2}

\section{Figures}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/signs-language-recognition.png} 
\caption{Signs Language Recognition}
\label{fig-slr}
\end{figure}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/sigml-example.png} 
\caption{SiGML transcription of 'here'}
\label{fig-sigml}
\end{figure}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/temporal-modeling.png} 
\caption{Temporal modeling using SubUNets}
\label{temporal-modeling}
\end{figure}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/signing-space.png} 
\caption{Signing Space}
\label{fig-signing-space}
\end{figure}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/signing-locations.png} 
\caption{Signing Locations}
\label{fig-signing-locations}
\end{figure}

\begingroup
\footnotesize
\begin{landscape}
\begin{longtable}{|>{\hspace{0pt}}m{0.142\linewidth}|>{\hspace{0pt}}m{0.083\linewidth}|>{\hspace{0pt}}m{0.065\linewidth}|>{\hspace{0pt}}m{0.063\linewidth}|>{\hspace{0pt}}m{0.077\linewidth}|>{\hspace{0pt}}m{0.063\linewidth}|>{\hspace{0pt}}m{0.109\linewidth}|>{\hspace{0pt}}m{0.073\linewidth}|>{\hspace{0pt}}m{0.106\linewidth}|>{\hspace{0pt}}m{0.079\linewidth}|>{\hspace{0pt}}m{0.063\linewidth}|}
\caption{Public Available Large Scale Datasets}\\ 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} DATASET & LANG & SIG\par{}NERS & CLA\par{}SSES & VIDEO \par{}INST. & DUR.~\par{}(H) & RES. & FPS & TYPE & MODA\par{}LITIES & YEAR \endfirsthead 
\hline
Signum SI & German & 25 & 780 & 19500 & 55.3 & 776x578 & 30 & continuous & RGB & 2007 \\ 
\hline
Signum isol. & German & 25 & 455 & 11375 & 8.43 & 776x578 & 30 & both & RGB & 2007 \\ 
\hline
Signum subset & German & 1 & 780 & 2340 & 4.92 & 776x578 & 30 & both & RGB & 2007 \\ 
\hline
Phoenix SD & German & 9 & 1231 & 6841 & 10.71 & 210x260 & 25 & continuous & RGB & 2014 \\ 
\hline
Phoenix SI & German & 9 & 1117 & 4667 & 7.28 & 210x260 & 25 & continuous & RGB & 2014 \\ 
\hline
CSL SD & Chinese & 50 & 178 & 25000 & 100+ & 1920x1080 & 30 & continuous & RGB+D & 2016 \\ 
\hline
CSL SI & Chinese & 50 & 178 & 25000 & 100+ & 1920x1080 & 30 & continuous & RGB+D & 2016 \\ 
\hline
CSL isol. & Chinese & 50 & 500 & 125000 & 67.75 & 1920x1080 & 30 & isolated & RGB+D & 2016 \\ 
\hline
Phoenix-T & German & 9 & 1231 & 8257 & 10.53 & 210x260 & 25 & continuous & RGB & 2018 \\ 
\hline
ASL 100 & English & 189 & 100 & 5736 & 5.55 & varying & varying & isolated & RGB & 2019 \\ 
\hline
ASL 1000 & English & 222 & 1000 & 25513 & 24.65 & varying & varying & isolated & RGB & 2019 \\ 
\hline
GSL isol. & Greek & 7 & 310 & 40785 & 6.44 & 848x480 & 30 & isolated & RGB+D & 2019 \\ 
\hline
GSL SD & Greek & 7 & 310 & 10290 & 9.59 & 848x480 & 30 & continuous & RGB+D & 2019 \\
\hline
\end{longtable}
\label{tab:public-large-datasets}
\end{landscape}

\endgroup