### 2021-05-09 phd/thesis plan/Tables/researchbudget.tex

\begingroup
\footnotesize
\begin{longtable}{|>{\hspace{0pt}}m{0.286\linewidth}|>{\hspace{0pt}}m{0.357\linewidth}|>{\hspace{0pt}}m{0.104\linewidth}|>{\hspace{0pt}}m{0.19\linewidth}|} 
\hline
\rowcolor[rgb]{0.502,0.502,0.502} RESOURCES & DESCRIPTION & COST & FUNDING \endfirsthead 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} SOFTWARE LICENSING &  & USD 0.00 &  \\ 
\hline
Pycharm (Student license) & IDE for python development &  &  \\ 
\hline
AppCode (Student license) & IDE for Apps development &  &  \\ 
\hline
CLion (Student license) & IDE for C and C++ development &  &  \\ 
\hline
iOS & Mobile operating system &  &  \\ 
\hline
MacOSx & Desktop operating system &  &  \\ 
\hline
TextMaker & LateX editor &  &  \\ 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} CLOUD SERVICES &  & USD 5508.00 &  \\ 
\hline
AWS Sage Maker & \textcolor[rgb]{0.086,0.098,0.122}{5 jobs per month x 1 instances }\par{}\textcolor[rgb]{0.086,0.098,0.122}{per job x 8 hours per job = 40.00 }\par{}\textcolor[rgb]{0.086,0.098,0.122}{SageMaker Training hours per month}\par{}\textcolor[rgb]{0.086,0.098,0.122}{}\textcolor[rgb]{0.086,0.098,0.122}{}\par{}\textcolor[rgb]{0.086,0.098,0.122}{40.00 hours per month x 3.825 USD }\par{}\textcolor[rgb]{0.086,0.098,0.122}{per hour instance cost = 153.00 USD }\par{}\textcolor[rgb]{0.086,0.098,0.122}{(monthly On-Demand cost)}\par{}\textcolor[rgb]{0.086,0.098,0.122}{}\textbf{Total cost for SageMaker Training (monthly): }\par{}\textbf{153.00 USD}\textcolor[rgb]{0.086,0.098,0.122}{}\textcolor[rgb]{0.086,0.098,0.122}{} & USD 5508.00 & \par{}Student Education Bonus\par{}provided by employer.\par{} \\ 
\hline
GitHub & Source code versioning and control & USD 0.00 &  \\ 
\hline
S3 (Student version) & Cloud objects storage & USD 0.00 &  \\ 
\hline
Lambda / Serverless (Student version) & Functions as a service & USD 0.00 &  \\ 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} PERSONEL &  & USD 0.00 &  \\ 
\hline
Doctoral student &  & USD 0.00 &  \\ 
\hline
Doctoral advisor &  & USD 0.00 &  \\ 
\hline
\end{longtable}
\endgroup