### 2021-05-09 phd/thesis plan/Tables/targetindexedjournals.tex

\begingroup
\footnotesize
\begin{longtable}{|l|l|l|}
\caption{List of target indexed journals}\\ 
\hline
\rowcolor[rgb]{0.502,0.502,0.502} NAME & OVERVIEW & URL \endfirsthead 
\hline
\begin{tabular}[c]{@{}l@{}}International Journal \\of Machine Learning \\and Cybernetics\end{tabular} & \begin{tabular}[c]{@{}l@{}}The International Journal of Machine Learning\\~and Cybernetics (IJMLC) focuses on the key \\research problems emerging at the junction of \\machine learning and cybernetics and serves\\~as a broad forum for rapid dissemination of the \\latest advancements in the area. The emphasis \\of IJMLC is on the hybrid development of \\machine learning and cybernetics schemes \\inspired by different contributing disciplines \\such as engineering, mathematics, cognitive \\sciences, and applications. New ideas, design \\alternatives, implementations and case studies \\pertaining to all the aspects of machine learning\\~and cybernetics fall within the scope of the IJMLC.\end{tabular} & \begin{tabular}[c]{@{}l@{}}https://www.springer\\.com/journal/\\13042\end{tabular} \\ 
\hline
\begin{tabular}[c]{@{}l@{}}Machine Learning \\with Applications\end{tabular} & \begin{tabular}[c]{@{}l@{}}Machine Learning with Applications (MLWA) is a \\peer reviewed, open access journal focused on \\research related to machine learning. The journal \\encompasses all aspects of research and \\development in ML, including but not limited \\to data mining, computer vision, natural language \\processing (NLP), intelligent systems, neural networks, \\AI-based software engineering, bioinformatics and their \\applications in the areas of engineering, medicine, \\biology, education, business and social sciences. \\It covers a broad spectrum of applications in the \\community, from industry, government, and academia.\end{tabular} & \begin{tabular}[c]{@{}l@{}}https://www.journals.\\elsevier.com/\\machine-learning\\-with-applications\end{tabular} \\ 
\hline
\begin{tabular}[c]{@{}l@{}}\textcolor[rgb]{0.133,0.133,0.133}{Computers \& Electrical }\\\textcolor[rgb]{0.133,0.133,0.133}{Engineering}\end{tabular} & \begin{tabular}[c]{@{}l@{}}Published since 1973, Computers \& Electrical \\Engineering provides rapid publication of topical \\research into the integration of computer technology \\and computational techniques with electrical and \\electronic systems. The journal publishes papers \\featuring novel implementations of computers and \\computational techniques in areas like signal \\processing, high-performance computing, \\artificial intelligence, and communications.\end{tabular} & \begin{tabular}[c]{@{}l@{}}https://www.journals.\\elsevier.com/\\computers-and-\\electrical-engineering\end{tabular} \\
\hline
\end{longtable}

\endgroup